Optimal Algorithms for Dynamic Balanced Graph Partitioning

### 2020-05-13 paper/disc20clustering.tex
special => pivot node

We start by constructing a~ground set of size $k-1$ on an arbitrarily chosen cluster.
In any partition, there must exist an isolated node collocated with the ground set of size $k-1$, and we call it \emph{special}.
We issue requests between the current special node, and some node that was initially collocated with it.
By repeating such requests, almost every node is collocated with the first ground set at some point .
In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two node exchanges ("swaps").
Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.

@@

\begin{proof}
	Fix any online algorithm \ALG.
	For a ground set $C$ of nodes that are initially collocated in one cluster,
	let $I(C)$ denote the cluster,
	i.e., the cluster that is the \emph{origin} of $C$.
%	If $C$ contains nodes that originate from different clusters
%	then $I(C)$ is undefined.
%	Throughout the construction, we maintain ground sets for each node.
	Initially, all nodes are isolated, i.e., each node is in a singleton ground set.
	First, we choose a cluster arbitrarily and
	 create a~ground set $B$ of $k-1$ nodes in this cluster.
	%We never dismiss any constructed ground set.
	Each cluster hosts exactly $k$ nodes, and in any feasible partition,
	 a~single isolated node must be collocated with $B$.
	At any time,
	we refer to the isolated node currently collocated with $B$ as the \emph{special} node.
	Let $x_0$ denote the first special node.

	Then, we join the special node to a~larger ground set to force its eviction.
	Precisely,
	we create a~ground set $\{x_0, y_0\}$, 	where
%	 $x_0$ is the current special node and
	 $y_0$ is an arbitrary isolated node.
	Since \ALG does not have $\{x_0,y_0\}$ collocated, 
	we issue an external request between them so that \ALG collocates them.
	\ALG cannot collocate $\{x_0, y_0\}$ with $B$ (as $B$'s size is $k-1$), hence it collocates them in a different cluster.
	In order to preserve a~feasible partition of nodes after collocating $\{x_0, y_0\}$,
	\ALG must replace $x_0$ with another isolated node.
%	denoted by $x_1$.

	We proceed by joining the current special node to a~larger ground set in a similar way,
	as long as a feasible partition exists afterwards.
	Consider the isolated node $x_i$ currently collocated with $B$.
	 We issue a request between $x_i$ and some node $y_i \in C_i$,
	 where $C_i$ is the largest ground set
	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$,
	 only if  a feasible partition exists
	 that collocates all nodes in $\{x_i\} \cup C_i $.
	Any such feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
	Otherwise,
	no such feasible partition exists and
	we terminate the process.

@@

	\ALG performs at least one swap each time a~ground set grows,
	i.e., each time it joins the isolated node $x_i$ (collocated with $B$), to  the ground set $C_i$ as specified before.
%	In other words,
%	some ground set grows by one node at each of these steps.
	Consider any ground set $C^*$ after serving $\sigma$.
	This ground set has grown exactly $|C^*|-1$ times until the end of $\sigma$. 
	Let $\mathcal{C}^*$ be the set of all ground sets after serving the last request in $\sigma$.
	Thus, $\mathcal{C}^*$ includes ground sets $B$,
	$\{x_0,y_0\}$,
	and $\ell+1$ singleton ground sets.
%	(including $x^*$ and possibly $y^*$).	
	Each time we join the special node to the largest ground set of same origin.
	Hence, every  other ground set in $\mathcal{C}^*$ has grown maximally.
	There are at most $\ell-1$ such ground sets in addition to the ones 
	we considered so far,
	one per cluster except the cluster containing $B$. 
	Therefore,  $|\mathcal{C}^*| \leq 1 + 1 + (\ell+1)  + (\ell-1) = 2\ell+2$.
	Note that among all non-singleton ground sets in $\mathcal{C}^*$,
	 only $B$ does not grow during the process.
	 Thus,
    the total number of times that 
	a ground set in $\mathcal{C}^*$  has grown is
	\begin{align*}
	\sum_{C^* \in \mathcal{C}^* } (|C^*|-1) - (k-1) &= 
	\sum_{C^* \in \mathcal{C}^* } |C^*| - 
		\sum_{C^* \in \mathcal{C}^* } 1 - (k-1) \\
	&\geq  k  \ell - (2\ell+2) - (k-1) =  (k-2)(\ell-1) - 3,
	\end{align*}
	which is the least number of swaps performed by \ALG.
	The competitive ratio is then
	$\ALG / \OPT \geq ((k-2)(\ell-1) - 3) / 2$.
	%%
	\begin{comment}
	\ALG performs at least one swap  every time   a~ground set is constructed.
	Each expansion reduces the number of isolated nodes by $1$ or $2$.
	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
	\end{comment}
	%%
\end{proof}

>>>>

We start by constructing a~ground set of size $k-1$ on an arbitrarily chosen cluster.
In any partition, there must exist an isolated node collocated with the ground set of size $k-1$.
%, and we refer to it as the \emph{pivot} node.
We issue requests between this node and some node that was initially collocated with it.
By repeating such requests, almost every node is once collocated with the first ground set.
In comparison, we show that there exists an optimal offline algorithm \OPT that performs only two node exchanges ("swaps").
Figure~\ref{fig:lb} illustrates the constructs used in the lower bound.

@@

\begin{proof}
	Fix any online algorithm \ALG.
	For a ground set $C$ of nodes that are initially collocated in one cluster,
	let $I(C)$ denote the cluster,
	i.e., the cluster that is the \emph{origin} of $C$.
%	If $C$ contains nodes that originate from different clusters
%	then $I(C)$ is undefined.
%	Throughout the construction, we maintain ground sets for each node.
	Initially, all nodes are isolated, i.e., each node is in a singleton ground set.
	First, we choose a cluster arbitrarily and
	 create a~ground set $B$ of $k-1$ nodes in this cluster.
	%We never dismiss any constructed ground set.
	Each cluster hosts exactly $k$ nodes, and in any feasible partition,
	 a~single isolated node must be collocated with $B$.
	At any time,
	we refer to the isolated node currently collocated with $B$ as the \emph{pivot} node.
	Let $x_0$ denote the first pivot node.

	Then, we join the pivot node to a~larger ground set to force its eviction.
	Precisely,
	we create a~ground set $\{x_0, y_0\}$, 	where
%	 $x_0$ is the current pivot node and
	 $y_0$ is an arbitrary isolated node.
	Since \ALG does not have $\{x_0,y_0\}$ collocated, 
	we issue an external request between them so that \ALG collocates them.
	\ALG cannot collocate $\{x_0, y_0\}$ with $B$ (as $B$'s size is $k-1$), hence it collocates them in a different cluster.
	In order to preserve a~feasible partition of nodes after collocating $\{x_0, y_0\}$,
	\ALG must replace $x_0$ with another isolated node.
%	denoted by $x_1$.

	We proceed by joining the current pivot node to a~larger ground set in a similar way,
	as long as a feasible partition exists afterwards.
	Consider the isolated node $x_i$ currently collocated with $B$.
	 We issue a request between $x_i$ and some node $y_i \in C_i$,
	 where $C_i$ is the largest ground set
	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$,
	 only if  a feasible partition exists
	 that collocates all nodes in $\{x_i\} \cup C_i $.
	Any such feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
	Otherwise,
	no such feasible partition exists and
	we terminate the process.

@@

	\ALG performs at least one swap each time a~ground set grows,
	i.e., each time it joins the isolated node $x_i$ (collocated with $B$), to  the ground set $C_i$ as specified before.
%	In other words,
%	some ground set grows by one node at each of these steps.
	Consider any ground set $C^*$ after serving $\sigma$.
	This ground set has grown exactly $|C^*|-1$ times until the end of $\sigma$. 
	Let $\mathcal{C}^*$ be the set of all ground sets after serving the last request in $\sigma$.
	Thus, $\mathcal{C}^*$ includes ground sets $B$,
	$\{x_0,y_0\}$,
	and $\ell+1$ singleton ground sets.
%	(including $x^*$ and possibly $y^*$).	
	Each time we join the pivot node to the largest ground set of same origin.
	Hence, every  other ground set in $\mathcal{C}^*$ has grown maximally.
	There are at most $\ell-1$ such ground sets in addition to the ones 
	we considered so far,
	one per cluster except the cluster containing $B$. 
	Therefore,  $|\mathcal{C}^*| \leq 1 + 1 + (\ell+1)  + (\ell-1) = 2\ell+2$.
	Note that among all non-singleton ground sets in $\mathcal{C}^*$,
	 only $B$ does not grow during the process.
	 Thus,
    the total number of times that 
	a ground set in $\mathcal{C}^*$  has grown is
	\begin{align*}
	\sum_{C^* \in \mathcal{C}^* } (|C^*|-1) - (k-1) &= 
	\sum_{C^* \in \mathcal{C}^* } |C^*| - 
		\sum_{C^* \in \mathcal{C}^* } 1 - (k-1) \\
	&\geq  k  \ell - (2\ell+2) - (k-1) =  (k-2)(\ell-1) - 3,
	\end{align*}
	which is the least number of swaps performed by \ALG.
	The competitive ratio is then
	$\ALG / \OPT \geq ((k-2)(\ell-1) - 3) / 2$.
	%%
	\begin{comment}
	\ALG performs at least one swap  every time   a~ground set is constructed.
	Each expansion reduces the number of isolated nodes by $1$ or $2$.
	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
	\end{comment}
	%%
\end{proof}
